---
title: "Quicktour"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quicktour}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> Port of the Quicktour vignette in 🤗 tokenizers website. Click [here](https://huggingface.co/docs/tokenizers/python/latest/quicktour.html#build-a-tokenizer-from-scratch) for the
original version.

```{r setup}
library(hftokenizers)
```

Let’s have a quick look at the 🤗 Tokenizers library features. The library provides an implementation of today’s most used tokenizers that is both easy to use and blazing fast.

It can be used to instantiate a pretrained tokenizer but we will start our quicktour by building one from scratch and see how we can train it.

## Build a tokenizer from scratch

To illustrate how fast the 🤗 Tokenizers library is, let’s train a new tokenizer on wikitext-103 (516M of text) in just a few seconds. First things first, you will need to download this dataset and unzip it with:

```{r}
url <- "https://s3.amazonaws.com/research.metamind.io/wikitext/wikitext-103-raw-v1.zip"
fpath <- pins::pin(url, name = "wikitext")
```
### Training the tokenizer

In this tour, we will build and train a Byte-Pair Encoding (BPE) tokenizer. For 
more information about the different type of tokenizers, check out this guide in 
the 🤗 Transformers documentation. Here, training the tokenizer means it will 
learn merge rules by:

- Start with all the characters present in the training corpus as tokens.
- Identify the most common pair of tokens and merge it into one token.
- Repeat until the vocabulary (e.g., the number of tokens) has reached the size we want.

The main API of the library is the class Tokenizer, here is how we instantiate 
one with a BPE model:

```{r}
tokenizer <- tokenizer$new(model = models_bpe$new(unk_token = "[UNK]"))
```

To train our tokenizer on the wikitext files, we will need to instantiate a trainer, 
in this case a `BpeTrainer`:

```{r}

```


