---
title: "Quicktour"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quicktour}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> Port of the Quicktour vignette in 🤗 tokenizers website. Click [here](https://huggingface.co/docs/tokenizers/python/latest/quicktour.html#build-a-tokenizer-from-scratch) for the
original version.

```{r setup}
library(hftokenizers)
```

Let’s have a quick look at the 🤗 Tokenizers library features. The library provides an implementation of today’s most used tokenizers that is both easy to use and blazing fast.

It can be used to instantiate a pretrained tokenizer but we will start our quicktour by building one from scratch and see how we can train it.

## Build a tokenizer from scratch

To illustrate how fast the 🤗 Tokenizers library is, let’s train a new tokenizer on wikitext-103 (516M of text) in just a few seconds. First things first, you will need to download this dataset and unzip it with:

```{r}
url <- "https://s3.amazonaws.com/research.metamind.io/wikitext/wikitext-103-raw-v1.zip"
fpath <- pins::pin(url)
```
### Training the tokenizer

In this tour, we will build and train a Byte-Pair Encoding (BPE) tokenizer. For 
more information about the different type of tokenizers, check out this guide in 
the 🤗 Transformers documentation. Here, training the tokenizer means it will 
learn merge rules by:

- Start with all the characters present in the training corpus as tokens.
- Identify the most common pair of tokens and merge it into one token.
- Repeat until the vocabulary (e.g., the number of tokens) has reached the size we want.

The main API of the library is the class Tokenizer, here is how we instantiate 
one with a BPE model:

```{r}
tokenizer <- tokenizer$new(model = models_bpe$new(unk_token = "[UNK]"))
```

To train our tokenizer on the wikitext files, we will need to instantiate a trainer, 
in this case a `BpeTrainer`:

```{r}
trainer <- trainers_bpe$new(
  special_tokens=c("[UNK]", "[CLS]", "[SEP]", "[PAD]", "[MASK]")
)
```

We can set the training arguments like `vocab_size` or `min_frequency` (here left 
at their default values of 30,000 and 0) but the most important part is to give 
the `special_tokens` we plan to use later on (they are not used at all during 
training) so that they get inserted in the vocabulary.

> The order in which you write the special tokens list matters: here "[UNK]" will 
  get the ID 0, "[CLS]" will get the ID 1 and so forth.
  
We could train our tokenizer right now, but it wouldn’t be optimal. Without a 
pre-tokenizer that will split our inputs into words, we might get tokens that 
overlap several words: for instance we could get an `"it is"` token since those 
two words often appear next to each other. Using a pre-tokenizer will ensure 
no token is bigger than a word returned by the pre-tokenizer. Here we want to 
train a subword BPE tokenizer, and we will use the easiest pre-tokenizer possible 
by splitting on whitespace.  

```{r}
tokenizer$pre_tokenizer <- pre_tokenizers_whitespace()
```

Now, we can just call the `train()` method with any list of files we want to use:

```{r}
tokenizer$train(fpath, trainer)
```

This should only take a few seconds to train our tokenizer on the full wikitext dataset! To save the tokenizer in one file that contains all its configuration and vocabulary, just use the `save()` method:

```{r}
tokenizer$save("tokenizer-wiki.json")
```

and you can reload your tokenizer from that file with the `from_file()` class method:

```{r}
tokenizer <- tokenizer$from_file("tokenizer-wiki.json")
```





